---
title: "Pregnancy loss predictors"
author: "Laura Symul"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::html_document2: 
    theme: flatly
    highlight: haddock
    toc: yes
    toc_float: true
    toc_depth: 5
    number_sections: true
    fig_caption: true
---

```{r preg_loss_pred setup , include = FALSE, eval = TRUE, cache = FALSE}
source("Scripts/00_setup.R")
```





# Predictors of pregnancy losses


## Computing statistics for the 4 cycles before 1st pregnancy tests

```{r preg_loss_pred stats for the 4 cycles before 1st pregnancy}

load(paste0(IO$output_data, "cycles.Rdata"), verbose = TRUE)


colnames(cycles)
cycles$cycle_nb_first_pos_preg_test = users$first_cycle_preg[match(cycles$user_id, users$user_id)]
cycles$cycles_before_preg_4 = (cycles$cycle_nb >= (cycles$cycle_nb_first_pos_preg_test - 4)) & 
  (cycles$cycle_nb <= (cycles$cycle_nb_first_pos_preg_test - 1))


users_cycles_stats = ddply(cycles[which(cycles$cycles_before_preg_4),], 
                            .(user_id), 
                            .fun = summarize,
                            avg_cycle_length_4 = mean(cycle_length, na.rm = TRUE),
                            median_cycle_length_4 = median(cycle_length, na.rm = TRUE),
                            sd_cycle_length_4 = sd(cycle_length, na.rm = TRUE)
)

head(users_cycles_stats)


column_names = colnames(users_cycles_stats)
column_names = column_names[-which(column_names %in% colnames(users))]
m = match(users$user_id, users_cycles_stats$user_id)
for(column  in column_names){
  eval(parse(text = paste0("users$",column,"= users_cycles_stats$",column,"[m]")))
  #eval(parse(text = paste0("users$",column,"[is.na(users$",column,")]= 0")))
}


# we also need to record the outcome of the first pregnancy

users_first_preg_outcome = cycles[which(cycles$cycle_nb == cycles$cycle_nb_first_pos_preg_test),c("user_id","preg_outcome")]

users$first_preg_outcome = users_first_preg_outcome$preg_outcome[match(users$user_id, users_first_preg_outcome$user_id)]

users$first_preg_outcome_simple = ifelse(users$first_preg_outcome %in% c("TB","BF","ExPTB","PTB"), "LB",
                                         ifelse(users$first_preg_outcome %in% c("EPL","LPL"),"PL",NA))

# and the time to first pregnancy test
users_date_first_preg_test = cycles[which(cycles$cycle_nb == cycles$cycle_nb_first_pos_preg_test),c("user_id","start_date")]

users$time_to_first_pos_test_in_days = as.numeric(
  users_date_first_preg_test$start_date[match(users$user_id,users_date_first_preg_test$user_id )] - 
    users$earliest_date)

users$time_to_first_pos_test = users$time_to_first_pos_test_in_days /365


#
save(users, file = paste0(IO$output_data,"users.Rdata"))
file.copy(from = paste0(IO$output_data,"users.Rdata"), to = paste0(IO$tmp_data,"users_with_stats_4_cycles.Rdata"), overwrite = TRUE)

```







## Is cycle length predictive of pregnancy outcomes

### stats on all previous cycles

```{r preg_loss_pred cycle length}

load(file = paste0(IO$output_data,"users.Rdata"), verbose = TRUE)

users$any_PL = ifelse(users$n_PL>0, TRUE, ifelse(users$n_LB>0,FALSE,NA))

u = users[which(!is.na(users$any_PL)),]


glm_cl = glm(
  any_PL ~ cycle_length_before_preg_avg + 
    cycle_length_before_preg_median + 
    cycle_length_before_preg_sd,
  data = u,
  family = "binomial")
summary(glm_cl)


glm_cl = glm(
  any_PL ~ cycle_length_before_preg_median,
  data = u,
  family = "binomial")

summary(glm_cl)


glm_cl = glm(
  any_PL ~ cycle_length_before_preg_sd,
  data = u,
  family = "binomial")

summary(glm_cl)

ggplot(u, aes(x = cycle_length_before_preg_avg, fill = any_PL))+
  geom_histogram(col = NA, alpha = 0.5, position = "identity", binwidth = 1)


ggplot(u, aes(x = cycle_length_before_preg_median, fill = any_PL))+
  geom_histogram(col = NA, alpha = 0.5, position = "identity", binwidth = 1)
  


ggplot(u, aes(x = cycle_length_before_preg_median, fill = any_PL))+
  geom_density(col = NA, alpha = 0.5, bw = 2)



ggplot(u, aes(x = log(cycle_length_before_preg_sd), fill = any_PL))+
  geom_histogram(col = NA, alpha = 0.5, position = "identity", binwidth = 1)

```


### stats on the 4 previous cycles

```{r preg_loss_pred cycle length - 4 cycles}

u = users[which(!is.na(users$first_preg_outcome_simple)),]
u$is_first_preg_a_PL = (u$first_preg_outcome_simple == "PL")

glm_cl = glm(
  is_first_preg_a_PL ~ avg_cycle_length_4 + 
    median_cycle_length_4 + 
    sd_cycle_length_4,
  data = u,
  family = "binomial")
summary(glm_cl)


glm_cl = glm(
  is_first_preg_a_PL ~ median_cycle_length_4,
  data = u,
  family = "binomial")

summary(glm_cl)


glm_cl = glm(
  is_first_preg_a_PL ~ sd_cycle_length_4,
  data = u,
  family = "binomial")

summary(glm_cl)


glm_cl = glm(
  is_first_preg_a_PL ~ 
    sd_cycle_length_4+
    median_cycle_length_4,
  data = u,
  family = "binomial")

summary(glm_cl)




ggplot(u, aes(x = avg_cycle_length_4, fill = is_first_preg_a_PL))+
  geom_histogram(col = NA, alpha = 0.5, position = "identity", binwidth = 1) + xlim(c(0,100))


ggplot(u, aes(x = median_cycle_length_4, fill = is_first_preg_a_PL))+
  geom_histogram(col = NA, alpha = 0.5, position = "identity", binwidth = 1) + xlim(c(0,100))


ggplot(u, aes(x = log10(sd_cycle_length_4), fill = is_first_preg_a_PL))+
  geom_histogram(col = NA, alpha = 0.5, position = "identity", binwidth = 0.1) 


ggplot(u, aes(x = sd_cycle_length_4, fill = is_first_preg_a_PL))+
  geom_histogram(col = NA, alpha = 0.5, position = "identity", binwidth = 1) + xlim(c(0,50))


```





## Is something else predictive of pregnancy outcomes?

### Detect ovulation in cycles without positive pregnancy tests

* Use HMM model (adapt it to include prior knowledge and LH tests)

* Label days table

* Aggregate cycles table with

    + ovulation day + uncertainty
  
    + temperature shift
  
* Aggregate users table with 

    + average, minimum and maximum temperature shift for cycles before first positive preg test




