---
title: "Reproductive objectives of the users"
author: "Laura Symul"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::html_document2: 
    theme: flatly
    highlight: haddock
    toc: yes
    toc_float: true
    toc_depth: 5
    number_sections: true
    fig_caption: true
---

```{r reprod_obj knitr options}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```



```{r reprod_obj setup, include = FALSE, eval = TRUE, cache = FALSE}
source("Scripts/00_setup.R")
```



# Reproductive objectives of the users 

## Data preparation

### Loading cycles and users

```{r reprod_obj load data}
users = read_feather(path = paste0(IO$output_data, "users.feather"))
cycles = read_feather(path = paste0(IO$output_data, "cycles.feather"))
pregnancies = read_feather(path = paste0(IO$output_data, "pregnancies.feather"))
```

### Identifying the 3 cycles preceding a pregnancy

```{r reprod_obj pregnancy id}

cycles$cycle_nb_from_next_preg = NA
j_preg = which(cycles$preg_test_class == "pregnant")
cycles$pregnancy_id = NA 
j3 = match(paste0(cycles$user_id[j_preg],"_",cycles$cycle_nb[j_preg] - 3),cycles$cycle_id)
cycles$cycle_nb_from_next_preg[j3] = -3 
cycles$pregnancy_id[j3[!is.na(j3)]] = cycles$cycle_id[j_preg[!is.na(j3)]] 
j2 = match(paste0(cycles$user_id[j_preg],"_",cycles$cycle_nb[j_preg] - 2),cycles$cycle_id)
cycles$cycle_nb_from_next_preg[j2] = -2
cycles$pregnancy_id[j2[!is.na(j2)]] = cycles$cycle_id[j_preg[!is.na(j2)]] 
j1 = match(paste0(cycles$user_id[j_preg],"_",cycles$cycle_nb[j_preg] - 1),cycles$cycle_id)
cycles$cycle_nb_from_next_preg[j1] = -1
cycles$pregnancy_id[j1[!is.na(j1)]] = cycles$cycle_id[j_preg[!is.na(j1)]] 
j0 = j_preg
cycles$cycle_nb_from_next_preg[j0] = 0
cycles$pregnancy_id[j0] = cycles$cycle_id[j_preg] 

```


## based on what they reported in the app at on-boarding

```{r reprod_obj pregnancy outcomes by reproductive objective as declared by the users}

ggplot(pregnancies, aes(x = preg_duration, fill = reprod_obj_app) )+
  geom_histogram(position = "identity",binwidth = 7, alpha = 1) +
  facet_grid(reprod_obj_app ~ . , scale = "free")+
  xlim(0,1000)


ggplot(pregnancies, aes(x = preg_outcome, fill = reprod_obj_app) )+
  geom_bar() +
  facet_grid(reprod_obj_app ~ . , scale = "free")

t1 = table( pregnancies$reprod_obj_app, pregnancies$preg_outcome)
round(100* t1/apply(t1, 1, sum))

```
It seems a little counter-intuitive as users who declared wanting to avoid pregnancy have the highest proportion of live-birth and the lowest proportion of pregnancy losses.

Maybe, it is because these labels are not reliable.

```{r reprod_obj reliability of the reproductive objective as declared by the users}

ggplot(cycles[cycles$n_insemination > 0,], aes(n_insemination, fill = reprod_obj_app))+
  geom_histogram(position = "identity",binwidth = 1) + xlim(0,20) +
  facet_grid(reprod_obj_app ~ ., scale = "free")

table(cycles$reprod_obj_app, cycles$n_insemination > 0)
table(cycles$reprod_obj_app, cycles$n_insemination > 0)/apply(table(cycles$reprod_obj_app, cycles$n_insemination > 0)
, 1, sum)



ggplot(cycles[cycles$n_prot_sex > 0,], aes(n_prot_sex, fill = reprod_obj_app))+
  geom_histogram(position = "identity",binwidth = 1) + xlim(0,20) +
  facet_grid(reprod_obj_app ~ ., scale = "free")

```

A lot of the users who declared wanting to avoid pregnancy are actually logging inseminations; similarly, many of those who declare wanting to achieve pregnancy log protected sexual intercourses.

It may thus be relevant to attempt to guess the reproductive objectives based on the sexual behavior of the users.

However, sexual intercourses are not logged in many cycles, including those in which positive pregnancy tests are logged.

```{r reprod_obj number of cycles in  which no sexual intercourses are logged}

table(cycles$n_tot_sex == 0) / sum(table(cycles$n_tot_sex == 0))

t2 = table(cycles$preg_test_class, cycles$n_tot_sex == 0)
t2
t2/apply(t2, 1, sum)

```

We can, nonetheless attempt to estimate the reproductive objectives for the users who do log sexual intercourses.

## based on their sexual behavior in the 3 previous cycles

We look at the overlap between their unprotected sexual intercourse and their fertility window in the 3 cycles preceding a pregnancy.


### Fertile window: counting backward

We can define the fertile window counting backward from the end of each cycle, assuming an identical luteal phase for each cycle and each users.

```{r reprod_obj adding fertile window to the days table}

input_days_folder = paste0(IO$output_data,"Days/")
tmp_folder = paste0(IO$tmp_data, "Days_with_fertility/")
if(!dir.exists(tmp_folder)){dir.create(tmp_folder)}

cl = makeCluster(par$n_cores)
registerDoParallel(cl)

days_files = list.files(input_days_folder)

foreach(file  = days_files, .packages = c('plyr','dplyr','feather')) %dopar%
{
  days = read_feather(path = paste0(input_days_folder,file))
  
  j = which(cycles$cycle_nb_from_next_preg %in% -3:-1)
  days = days[days$cycle_id %in% cycles$cycle_id[j],]
  
  days$fertility_counting = dict$fertility_counting$fertility[
    match(days$cycleday_from_end, dict$fertility_counting$cycleday_from_end)]
  days$fertility_counting[is.na(days$fertility_counting)] = 0
  
  cycles_agg = aggregate(fertility_counting ~ cycle_id, days, mean, na.rm = TRUE)
  
  days_mean_fertility = cycles_agg$fertility_counting[match(days$cycle_id, cycles_agg$cycle_id)]
  
  days$fertility_counting_n = days$fertility_counting - days_mean_fertility
  
  ok = write_feather(days, path = paste0(tmp_folder,file))
}

stopImplicitCluster()

```


### Reproductive objective scores per cycle

```{r reprod_obj reproductive objective score per cycle}

reproductive_obj_score = function(fertility, fertility_n, sex){
  score = sum(fertility_n * (sex == 2)) +
    -1 * sum(fertility * (sex == 1)) +
    -0.5 * sum(fertility * (sex == 3)) +
    2 * any(sex == 4)
  
  return(score)
}

tmp_folder = paste0(IO$tmp_data, "Days_with_fertility/")
days_files = list.files(tmp_folder)

cl = makeCluster(par$n_cores)
registerDoParallel(cl)

cycles_agg = foreach(file  = days_files, .combine = rbind, .packages = c('plyr','dplyr','feather')) %dopar%
{
  days = read_feather(path = paste0(tmp_folder,file))
  
  cycles_agg = ddply(days,
                     "cycle_id",
                     summarise,
                     reprod_obj_score = reproductive_obj_score(fertility_counting, fertility_counting_n, sex))
  
  return(cycles_agg)
}

stopImplicitCluster()

cycles$reprod_obj_score_counting = cycles_agg$reprod_obj_score[match(cycles$cycle_id, cycles_agg$cycle_id)]

write_feather(cycles, path = paste0(IO$output_data, "cycles.feather"))
ok = file.copy(from = paste0(IO$output_data, "cycles.feather") , to = paste0(IO$tmp_data, "cycles_with_reprod_obj_scores.feather"), overwrite = TRUE)

```


### Reproductive objective scores per pregnancy

```{r reprod_obj reproductive objective per pregnancy}

score_agg = aggregate(reprod_obj_score_counting ~  pregnancy_id  , cycles, median, na.rm = TRUE)
pregnancies$reprod_obj_score_counting = score_agg$reprod_obj_score_counting[match(pregnancies$pregnancy_id, score_agg$pregnancy_id)]

sex_agg = aggregate(n_tot_sex ~ pregnancy_id, cycles, sum, na.rm = TRUE)
pregnancies$n_tot_sex = sex_agg$n_tot_sex[match(pregnancies$pregnancy_id, sex_agg$pregnancy_id)]


pregnancies$reprod_obj_counting = cut(pregnancies$reprod_obj_score_counting, 
                             breaks = c(-Inf, -0.1,0.1,Inf),
                             labels = c("avoid_preg","unknown","get_preg"))


ggplot(pregnancies, aes(x = reprod_obj_score_counting, fill = reprod_obj_counting))+
  geom_histogram(binwidth = 0.1)

write_feather(pregnancies, path = paste0(IO$output_data,"pregnancies.feather"))
file.copy(from = paste0(IO$output_data,"pregnancies.feather"), to = paste0(IO$tmp_data,"pregnancies_with_3_cycles_score.feather") , overwrite = TRUE)

```


### Consistency between user-declared objectives and computed objectives

```{r reprod_obj consistency}

ggplot(pregnancies[which(pregnancies$reprod_obj_score_counting != 0),], 
       aes(x = reprod_obj_score_counting, fill = reprod_obj_counting))+
  geom_histogram(position = "identity", binwidth = 0.1, alpha = 1) +
  facet_grid(reprod_obj_app ~ . , scale = "free")+
  geom_vline(xintercept = 0)+
  xlim(c(-3,3))


```



### Pregnancy outcomes per reproductive objective

```{r reprod_obj visualization of pregnancy duration per reprod objectives}

ggplot(pregnancies, aes(x = preg_duration, fill = reprod_obj_counting) )+
  geom_histogram(position = "identity",binwidth = 7, alpha = 0.5) +
  facet_grid(reprod_obj_counting ~ . , scale = "free")+
  xlim(0,1000)


ggplot(pregnancies, aes(x = preg_outcome, fill = reprod_obj_counting) )+
  geom_bar() +
  facet_grid(reprod_obj_counting ~ . , scale = "free")

```


Here again, it seems that the users who are trying to avoid pregnancy have proportionally less induced or spontaneous pregnancy losses.

## Based on the sexual behavior within the cycle at which the users got pregnant

The 3 previous cycles may not reflect the intentions of the user at the cycle at which they got pregnant, as they may have changed their mind and got pregnant quite rapidly.

But maybe, we can guess whether they were trying to avoid a pregnancy if they logged protected sex or withdrawal before they logged their first positive pregnancy test. If they were, it may mean that these users got pregnant from a mis-estimation of their fertile window.

```{r reprod_obj protected sex or withdrawal before first positive pregnancy test}


cycle_ids = cycles$cycle_id[!is.na(cycles$preg_outcome)]
input_days_folder = paste0(IO$output_data,"Days/")

tic()
cl = makeCluster(par$n_cores)
registerDoParallel(cl)

days_files = list.files(input_days_folder)

cycles_agg = foreach(file  = days_files, .combine = rbind, .packages = c('plyr','dplyr','feather')) %dopar%
{
  days = read_feather(path = paste0(input_days_folder,file))
  
  days$fluid_eggwhite[is.na(days$fluid_eggwhite)] = 0
  days$fluid_watery[is.na(days$fluid_watery)] = 0
  days$fertile_mucus = days$fluid_eggwhite + days$fluid_watery
  
  j = which(days$cycle_id %in% cycle_ids)
  days = days[j,]
  days_first_pos_preg_test = cycles$day_first_pos_preg_test[match(days$cycle_id, cycles$cycle_id)]
  j = which(days$cycleday < days_first_pos_preg_test)
  
  cycles_agg = ddply(days[j,],
                     "cycle_id",
                     summarise,
                     n_prot_sex_no_mucus = sum((sex == 1) & (fertile_mucus == 0), na.rm = TRUE),
                     n_withdrawal_no_mucus = sum((sex == 3) & (fertile_mucus == 0), na.rm = TRUE),
                     n_unprot_sex_no_mucus = sum((sex == 2) & (fertile_mucus == 0), na.rm = TRUE),
                     n_insemination_no_mucus = sum((sex == 4) & (fertile_mucus == 0), na.rm = TRUE),
                     n_prot_sex_mucus = sum((sex == 1) & (fertile_mucus > 0), na.rm = TRUE),
                     n_withdrawal_mucus = sum((sex == 3) & (fertile_mucus > 0), na.rm = TRUE),
                     n_unprot_sex_mucus = sum((sex == 2) & (fertile_mucus > 0), na.rm = TRUE),
                     n_insemination_mucus = sum((sex == 4) & (fertile_mucus > 0), na.rm = TRUE))
  
  return(cycles_agg)
}

stopImplicitCluster()
toc()

cycles_agg_long = melt(cycles_agg, id.vars = "cycle_id")
ggplot(cycles_agg_long, aes(x = value, fill = variable)) + geom_histogram() + facet_grid(variable ~ ., scale = "free") + xlim(c(0.2,30))


#table(pmin(5, cycles_agg$n_prot_sex), pmin(10, cycles_agg$n_unprot_sex))

```

```{r reprod_obj adding n_sex to pregnancies}
m = match(pregnancies$pregnancy_id, cycles_agg$cycle_id)

colnames_to_add = colnames(cycles_agg[,-1])
for(col_ in colnames_to_add){
  eval(parse(text = paste0("pregnancies$",col_," = cycles_agg$",col_,"[m]")))
}

```


```{r reprod_obj score from sexual behavior in the cycle they got pregnant}

pregnancies$reprod_obj_score_in_cycle = 0 + 
  - 2*pregnancies$n_prot_sex_mucus +
  - 1*pregnancies$n_prot_sex_no_mucus +
  - 1 * pregnancies$n_withdrawal_mucus +
  - 0.5 * pregnancies$n_withdrawal_no_mucus 

pregnancies$reprod_obj_score_in_cycle[pregnancies$n_unprot_sex_mucus > 1] = 2

pregnancies$reprod_obj_score_in_cycle[(pregnancies$n_insemination_no_mucus + pregnancies$n_insemination_mucus)>0] = 10

pregnancies$reprod_obj_score_in_cycle[is.na(pregnancies$reprod_obj_score_in_cycle)] = 0

pregnancies$reprod_obj_in_cycle = cut(pregnancies$reprod_obj_score_in_cycle, breaks = c(-Inf, -0.5, 0.5, Inf),
                                      labels = c("avoid_preg","unknown","get_preg"))

```



```{r reprod_obj consistency with user declared objectives}

ggplot(pregnancies[pregnancies$reprod_obj_score_in_cycle <0,], aes(x = reprod_obj_score_in_cycle, fill = reprod_obj_app))+
  geom_histogram(binwidth = 0.5)+
  facet_grid(reprod_obj_app ~. , scale = "free")

table(pregnancies$reprod_obj_in_cycle, pregnancies$reprod_obj_app)

```

```{r reprod_obj pregnancy outcomes based on in-cycle sexual behavior}

ggplot(pregnancies, aes(x = preg_duration, fill = reprod_obj_in_cycle) )+
  geom_histogram(position = "identity",binwidth = 7, alpha = 0.5) +
  facet_grid(reprod_obj_in_cycle ~ . , scale = "free")+
  xlim(0,1000)


ggplot(pregnancies, aes(x = preg_outcome, fill = reprod_obj_in_cycle) )+
  geom_bar() +
  facet_grid(reprod_obj_in_cycle ~ . , scale = "free")


```


## Combining all of these indicators

If a user declared not wanting to get pregnant and logged sexual intercourses in the 3 previous cycles and in the conception cycle that were consistent with this declared objective, we can assume that this user was indeed trying to avoid pregnancy.

If, on the other side of the spectrum, a user declared that they wanted to achieve pregnancy and were acting consistently or received insemination, we can safely assume that they wanted to get pregnant. 

In all other cases, we will label the user with an "unknown" reproductive objective for the specific pregnancies.

```{r reprod_obj combining the 3 indicators}

pregnancies$reprod_obj = "unknown"

j = which((pregnancies$reprod_obj_app == "avoid_preg") & 
            ((pregnancies$reprod_obj_counting == "avoid_preg") | 
               (is.na(pregnancies$reprod_obj_counting) & (pregnancies$cycle_nb < 4))) & 
            (pregnancies$reprod_obj_in_cycle %in% c("avoid_preg","unknown")))
pregnancies$reprod_obj[j] = "avoid_preg"


j = which((pregnancies$reprod_obj_app %in% c("get_preg","preg")) & 
            ((pregnancies$reprod_obj_counting == "get_preg") | 
               (is.na(pregnancies$reprod_obj_counting) & (pregnancies$cycle_nb < 4))) & 
            (pregnancies$reprod_obj_in_cycle %in% c("get_preg","unknown")))
pregnancies$reprod_obj[j] = "get_preg"

table(pregnancies$reprod_obj)
table(pregnancies$reprod_obj_app)
table(pregnancies$reprod_obj_counting)
table(pregnancies$reprod_obj_in_cycle)


write_feather(pregnancies, path = paste0(IO$output_data,"pregnancies.feather"))
file.copy(from = paste0(IO$output_data,"pregnancies.feather"), to = paste0(IO$tmp_data,"pregnancies_with_reprod_obj.feather") , overwrite = TRUE)




```

```{r reprod_obj add reprod_obj at first pregnancy to the user table}

users$reprod_obj_at_first_pregnancy = pregnancies$reprod_obj[match(paste0(users$user_id, "_",users$first_cycle_preg),pregnancies$pregnancy_id)]


table(users$reprod_obj_at_first_pregnancy, users$reprod_obj_app)

write_feather(users, path = paste0(IO$output_data,"users.feather"))
file.copy(from = paste0(IO$output_data,"users.feather"), to = paste0(IO$tmp_data,"users_with_reprod_obj_at_first_pregnancy.feather"), overwrite = TRUE)

```


### Pregnancy outcomes per reproductive objectives (combined indicator)

```{r reprod_obj pregnancy outcomes per reprod obj - combined indicator}

ggplot(pregnancies, aes(x = preg_duration, fill = reprod_obj) )+
  geom_histogram(position = "identity",binwidth = 7, alpha = 0.5) +
  facet_grid(reprod_obj ~ . , scale = "free")+
  xlim(0,1000)


ggplot(pregnancies, aes(x = preg_outcome, fill = reprod_obj) )+
  geom_bar() +
  facet_grid(reprod_obj ~ . , scale = "free")


```

